---
title: "data_int"
author: "Weihao Li"
date: "2020-03-23"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load libraries

```{r 'load_packages', warning = FALSE, message = FALSE}
library(tidyverse)
library(sf)
library(furrr)
library(lubridate)
library(rnaturalearth)
library(gganimate)
library(bomrang)
```

# data summary

- [x] hostpots data 
    - [x] from 2015-07 to now
- [ ] weather data
    - [x] daily rainfall
    - [x] daily temperature
    - [x] daily solar exposure
    - [ ] lightning data
    - [ ] wind data
- [ ] road maps
    - [x] VIC main road maps
    - [ ] minor road maps
- [ ] recreation site
    - [x] VIC recreation site
- [ ] CFA
    - [x] VIC CFA stations location
    - [ ] CFA bushfire (not up to date)


# hotspot data

data serverd on [P-Tree system](ftp://ftp.ptree.jaxa.jp/pub/himawari/L3/WLF/bet)
From 2015-07 to now

```{r 'define_read_hotspot'}
read_hotspot <- function(directory = "data/Himawari-8", from = '20190501', to = '20190701'){
  ### include from, exclude to
  
  fls = list.files(directory, full.names = TRUE)
  fls_short = list.files("data/Himawari-8", full.names = FALSE)
  fls_filter = (fls_short > paste('H08', from, sep = '_')) & (fls_short < paste('H08', to, sep = '_'))
  fls = fls[fls_filter]
  all = map(fls, read_csv, quote="'")
  d = bind_rows(all)
  return(d)
  
}
```

```{r 'read_hotspots', warning = FALSE, message = FALSE}
hotspots = read_hotspot(directory = "data/Himawari-8", from = '20190501', to = '20190701')
```

```{r 'get_au_hotspots'}
# apply australia coordinates 
hotspots_au = hotspots %>%
  filter(between(lon, 112, 155)) %>% 
  filter(between(lat, -44, -10))

```

```{r 'fire_power_filter'}

firepower_threshold = 100 #threshold for firepower

hotspots_au = hotspots_au %>%
  filter(firepower > firepower_threshold)

```

```{r 'hotspots_example'}

au_map = ne_states(country = 'Australia', returnclass = 'sf') #get au map
samples = select(hotspots_au, `#obstime`, lon, lat, firepower)

p = ggplot(samples) +
  geom_sf(data = au_map) +
  geom_point(aes(group = `#obstime`, x = lon, y = lat, size = firepower), color = 'red', alpha = 0.2) + 
  transition_time(`#obstime`) +
  exit_fade() +
  enter_fade() +
  labs(title = 'Time: {frame_time}') +
  theme_bw()

animate(p)

```

# Weather data

```{r 'get_stations_ids'}
# get all stations and sorted by distance from MEL
station_ids = sweep_for_stations(latlon = c(-37.8136, 144.9631))
```

```{r 'filter_VIC_stations'}
# filter VIC stations
VIC_station_ids = station_ids %>%
  filter(state == 'VIC')
```

```{r 'plot_stations_in_map'}
ggplot(VIC_station_ids) +
  geom_sf(data = au_map, fill = "grey", colour = "black")+
  coord_sf(ylim = c(-40,-34), xlim = c(140,150)) +
  geom_point(aes(x = lon, y = lat), color = 'blue') +
  theme_bw() +
  ggtitle('VIC weather stations')
```

```{r 'define_get_historical_weather'}

get_comprehensive_historical_weather <- function(station_id){

# requst weather data
rain_record = get_historical(station_id, type = "rain")
min_record = get_historical(station_id, type = "min")
max_record = get_historical(station_id, type = "max")
solar_record = get_historical(station_id, type = "solar")

# keep useful information
rain_record = select(rain_record, -quality, -product_code)
min_record = select(min_record, -quality, -product_code)
max_record = select(max_record, -quality, -product_code)
solar_record = select(solar_record, -product_code)

# combine data
t = full_join(rain_record, max_record, by = c('year','month','day','station_number'))
t = full_join(t, min_record, by = c('year','month','day','station_number'))
t = full_join(t, solar_record, by = c('year','month','day','station_number'))

return(t)
}


```

```{r 'get_historical_weather_example'}

weather_ex = get_comprehensive_historical_weather(VIC_station_ids$site[1])
weather_ex$date = as_date(paste(weather_ex$year,weather_ex$month,weather_ex$day,sep = '-'))
```

```{r 'show_weather_data'}
ggplot(filter(weather_ex, !is.na(rainfall))) +
  geom_line(aes(x = yday(date), y = rainfall)) +
  facet_wrap(~year)
```

```{r 'forecast_data'}
# used for finding the nearest forecast town
towns <- sweep_for_forecast_towns(latlon = c(-37.8136, 144.9631)) 
```

```{r 'towns_filter'}
VIC_towns <- towns %>%
  rowwise() %>%
  filter(grepl('VIC', aac))
```

```{r 'plot_forecast_towns'}
ggplot(VIC_towns) +
  geom_sf(data = au_map, fill = "grey", colour = "black")+
  coord_sf(ylim = c(-40,-34), xlim = c(140,150)) +
  geom_point(aes(x = lon, y = lat), color = 'blue') +
  theme_bw() +
  ggtitle('VIC forecast towns')
  
```

```{r 'get_forecast_data'}
# you will get a dataframe with 7-day weather forecast in the future
VIC_forecast = get_precis_forecast(state = 'VIC')
```


# Road data

[VIC_road_geometry](https://discover.data.vic.gov.au/dataset/victorian-road-geometry)

```{r 'define_read_VIC_road_map'}
read_VIC_road_map = function(directory = 'data/road_geometry'){
  fls = list.files(directory, full.names = TRUE)
  all = map(fls, read_csv, quote="'")
  for (i in seq(length(all))){
    all[[i]] <- all[[i]] %>%
      mutate(DIRECTION = as.character(DIRECTION))
  }
  d = bind_rows(all)
  return(d)
}
```


```{r 'read_VIC_road_map', message=FALSE, warning=FALSE}
VIC_road_map = read_VIC_road_map()
```

```{r 'show_road_map'}
# main road only

ggplot(VIC_road_map) +
  geom_sf(data = au_map, fill = "grey", colour = "black")+
  coord_sf(ylim = c(-40,-34), xlim = c(140,150)) +
  geom_segment(aes(x = S_LONGITUDE, y = S_LATITUDE, xend = E_LONGITUDE, yend = E_LATITUDE), color = 'blue') +
  theme_bw() +
  ggtitle('VIC road map')
```

# CFA

[VIC CFA stations location](https://discover.data.vic.gov.au/dataset/cfa-fire-station-vmfeat-geomark_point)

```{r 'read_cfa_position_map'}
shape = st_read("SDM697752/ll_gda94/sde_shape/whole/VIC/VMFEAT/layer/geomark_point.shp")
shape = shape %>% filter(FEATSUBTYP == "fire station")
```

```{r 'extract_stations_coordinates'}
stations = shape %>% 
  mutate(lon = st_coordinates(shape)[, 1],
         lat = st_coordinates(shape)[, 2]) %>%
  dplyr::select(name = NAME, lon, lat) %>%
  filter(!is.na(name))

```


```{r 'show_fire_stations_map'}
ggplot(stations) +
  geom_sf(data = au_map, fill = "grey", colour = "black")+
  coord_sf(ylim = c(-40,-34), xlim = c(140,150)) +
  geom_point(aes(x = lon, y = lat), color = 'blue', size =0.3)  +
  theme_bw() +
  ggtitle('VIC fire stations')
```

# Recreation Site

[VIC recreation site location](https://discover.data.vic.gov.au/dataset/recreation-sites)

```{r 'read_recreation_site_data'}
recreation = st_read("SDM698290\\ll_gda94\\sde_shape\\whole\\VIC\\FORESTS\\layer\\recweb_site.shp")

```

```{r 'show_recreation_map'}
ggplot(recreation) +
  geom_sf(data = au_map, fill = "grey", colour = "black")+
  coord_sf(ylim = c(-40,-34), xlim = c(140,150)) +
  geom_point(aes(x = LONGITUDE, y = LATITUDE), fill ='red', color = 'white', pch=21, alpha = 0.5, size = 3)+
  theme_bw() +
  ggtitle('VIC recreation site')
  
```

# bushfire cases

```{r 'read_fire_history'}
fire_history = st_read("SDM697447\\ll_gda94\\sde_shape\\whole\\VIC\\FIRE\\layer\\fire_history.shp")
bushfire_history = fire_history %>% 
  filter(FIRETYPE == 'BUSHFIRE')
bushfire_history = as.data.frame(bushfire_history)
```

```{r 'bushfire_data_processing'}
bushfire_history = arrange(bushfire_history, desc(START_DATE))
bushfire_location = bushfire_history$geometry
```

```{r 'show_fire_map'}
# many obs are from the same bushfire
ggplot() +
  geom_sf(data = au_map, fill = "grey", colour = "black")+
  geom_sf(data = bushfire_location[1:1000], fill = 'red', color = 'black') +
  coord_sf(ylim = c(-40,-34), xlim = c(140,150)) +
  theme_bw() +
  ggtitle('VIC last 1000 bushfire obs (last one: 2019-05-07)')
```



