---
title: 'Using Remote Sensing Data to Understand Fire Ignitions in Victoria during the 2019-2020 Australian Bushfire Season'
degreetype: 'Bachelor of Commerce (Honours)'
author: 'Weihao Li'
studentid: 28723740
output: 
  bookdown::pdf_book:
    toc: false
site: bookdown::bookdown_site
link-citations: yes
knit: "bookdown::render_book"
---

<!-- 
Edit these lines as appropriate.
The actual thesis content is in several Rmd files.

You'll need to edit the _bookdown.yml file to set the order in which you'd like them to appear. 

If you have specific LaTeX packages to add, put them in monashthesis.tex.

You will need to ensure you have installed the knitr and bookdown packages for R.

You will also need LaTeX installed on your computer.
-->

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r Prepare-Hotspots-For-Clustering, include=FALSE}
# Prepare hotspot data for the clustering algorithm
# Runtime: 5 minutes
# Input: hotspot data
# Output: VIC_hotspots_before_clustering.csv, VIC_hotspots_raw.csv
if (any(!file.exists(c("data/VIC_hotspots_before_clustering.csv", 
                       "data/VIC_hotspots_raw.csv")))){
  system("Rscript scripts/Clustering/Clustering-Python-setup.R")
} 
```


```{r Clustering-Parameters-Tuning, include=FALSE}
# Tune the parameters for the clustering algorithm
# Runtime: 8-10 hours
# Input: VIC_hotspots_before_clustering.csv
# Output: clustering_grid.csv
if (!file.exists("data/clustering_grid.csv")){
  system("python scripts/Clustering/clustering_tune.py")
} 
```


```{r Choose-Optimal-Parameters-For-Clustering, include=FALSE}
# Choose the optimal parameters based on the tuning result
# Runtime: 20 seconds
# Input: clustering_grid.csv
# Output: setting.txt, clustering_tuning_1.jpeg, clustering_tuning_2.jpeg
if (FALSE){
  system("Rscript scripts/Clustering/clustering_tune_vis.R")
}
```


```{r Hotspots-Clustering, include=FALSE}
# Cluster hotspots into bushfires
# Runtime: 6 minutes
# Input: VIC_hotspots_before_clustering.csv
# Output: VIC_hotspots_after_clustering.csv, summary.txt
if (!file.exists("data/VIC_hotspots_after_clustering.csv")){
  system("python scripts/Clustering/main.py")
}
```

```{r Bushfires-Data-Integration-Hotspots, include=FALSE}
# Data integration for bushfires (hotspots)
# Runtime: 20 minutes
# Input: VIC_hotspots_raw.csv, VIC_hotspots_after_clustering.csv, forest data, weather data, camping site, CFA stations, road map
# Output: predict_x.csv
if (!file.exists("data/predict_x.csv")){
  system("Rscript scripts/Training/combine_current.R")
}
```

```{r Bushfires-Data-Integration-Simulation, include=FALSE}
# Data integration for bushfires (Simulation)
# Runtime: 3 hours
# Input: forest data, weather data, camping site, CFA stations, road map
# Output: predict_x_grid.csv
if (!file.exists("data/predict_x_grid.csv")){
  system("Rscript scripts/Training/combine_grid.R")
}
```


```{r Bushfires-Data-Integration-History, include=FALSE}
# Data integration for bushfires (History)
# Runtime: 2 hours
# Input: fire origins, forest data, weather data, camping site, CFA stations, road map
# Output: training.csv
if (!file.exists("data/training.csv")){
  system("Rscript scripts/Training/combine_his.R")
}
```

```{r EDA, include=FALSE}
# EDA on bushfire history
# Runtime: 10-20 minutes
# Input: training.csv, VIC_hotspots_raw.csv, VIC_hotspots_after_clustering.csv
# Output: Many plots
if (FALSE){
  system("Rscript scripts/Training/EDA.R")
}

```

```{r Random-Forest-Model, include=FALSE}
# Random forest
# Runtime: 1 hour
# Input: training.csv
# Output: RF_best_features.rds, RF_model.rds
if (any(!file.exists(c("data/RF_best_features.rds", 
                       "data/RF_model.rds")))){
  system("Rscript scripts/Training/RF.R")
                       }
```

```{r MNL-Model, include=FALSE}
# MNL
# Runtime: 1 hour
# Input: training.csv
# Output: MNL_best_features.rds, MNL_model.rds
if (any(!file.exists(c("data/MNL_best_features.rds", 
                       "data/MNL_model.rds")))){
  system("Rscript scripts/Training/MNL.R")
                       }
```

```{r GAM-Model, include=FALSE}
# GAM
# Runtime: 6 hours
# Input: training.csv
# Output: GAM_best_features.rds, GAM_model.rds
if (any(!file.exists(c("data/GAM_best_features.rds", 
                       "data/GAM_model.rds")))){
  system("Rscript scripts/Training/GAM.R")
                       }
```

```{r XGB-Model, include=FALSE}
# XGB
# Runtime: 6 hours
# Input: training.csv
# Output: XGB_best_features.rds, XGB_model.rds
if (any(!file.exists(c("data/XGB_best_features.rds", 
                       "data/XGB_model.rds")))){
  system("Rscript scripts/Training/XGB.R")
                       }
```

```{r Final-Model, include = FALSE}
# Final Model
# Runtime: 2 minutes
# Input: RF_model.rds, MNL_model.rds, GAM_model.rds, XGB_model.rds, training.csv, RF_best_features.rds
# Output: Final_model.rds
if (!file.exists("data/Final_model.rds")){
  system("Rscript scripts/Training/Comparison.R")
                       }
```

```{r Prediction, include = FALSE}
# Prediction
# Runtime: 2 minutes
# Input: Final_model.rds, predict_x.csv, predict_x_grid.csv
# Output: prediction_2019-2020.csv, 2019-2020_prediction.jpeg, prediction_2019-2020_simulation.csv
if (!file.exists("data/prediction_2019-2020.csv")){
  system("Rscript scripts/Prediction/2019-2020_prediction.R")
}

if (!file.exists("data/prediction_2019-2020_simulation.csv")){
  system("Rscript scripts/Prediction/2019-2020_simulation_prediction.R")
                       }
```

# Acknowledgement {-}

I would like to thank my supervisors, Professor Dianne Cook and Emily Dodwell, for their invaluable support and guidance throughout this year. 

<!-- ## Software used {-} -->

`R 3.6.3` [@R], `RStudio 1.3.959` [@RStudio] and `Python 3.7.1` [@Python] are used for data analysis in this thesis. 

R packages used in this research include `tidyverse 1.3.0` [@R-tidyverse], `sf 0.9.5` [@R-sf], `lubridate 1.7.9` [@R-lubridate], `rnaturalearth 0.1.0` [@R-naturalearth], `here 0.1` [@R-here], `raster 3.3.13` [@R-raster], `bomrang 0.7.0` [@R-bomrang], `geodist 0.0.4` [@R-geodist], `progress 1.2.2` [@R-progress], `GGally 2.0.0` [@R-GGally], `naniar 0.5.2` [@R-naniar], `ggridges 0.5.2` [@R-ggridges], `ggthemes 4.2.0` [@R-ggthemes], `caret 6.0.86` [@R-caret], `lime 0.5.1` [@R-lime], `nnet 7.3.12` [@R-nnet], `randomForest 4.6.14` [@R-rf], `mgcv 1.8.31` [@R-mgcv], `fastDummies 1.6.2` [@R-fastDummies], `xgboost 1.1.1.1` [@R-xgboost], `plyr 1.8.6` [@R-plyr] and `pROC 1.16.2` [@R-pROC]. 

Python packages used in this research include `tqdm 4.48.2` [@Python-tqdm] and `numpy 1.19.1` [@Python-numpy]. 

This thesis is produced using the R package `bookdown 0.20` [@R-bookdown] with the R Markdown template for the honours thesis in the Department of Econometrics & Business Statistics, Monash University [@Monash-honours-template].

# Declaration {-}


I declare that this thesis contains no material which has been submitted in any form for the award of any other degree or diploma in any university or equivalent institution, and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

\makebox[0.5\linewidth][r]{}\dotsign\smallskip\
\hspace*{0.5\linewidth}Weihao Li\
\hspace*{0.5\linewidth}$4^{th}$ November 2020



<!-- anything we have done -> present tense -->
<!-- all others' past work, past events -> past tense -->
<!-- future work -> future tense -->



